---
title: "Week 06b - Measurement"
subtitle: "Bivariate Relationships: QQPlots and Matrices <html><div style='float:left'></div><hr color='#EB811B' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: "03/26/2020"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "macros.js"
      highlightStyle: github
      ratio: 4:3
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "http://cran.cnr.berkeley.edu/"
options(repos = r)
set.seed(12345)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #EB811B;
}

.orange {
  color: #EB811B;
}
</style>

# Today's Agenda

.font150[
* Recap: correlation

* Recap: log transformation

* QQ Plots

* Matrices
]

---

class: inverse, center, middle

# Recap: correlations

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---

# Recap

.font150[
* .orange[Z-Score:] number of standard deviations an observation is above or below the mean

* $z_{x_i} = \frac{x_i - \bar{x}}{S_x}$

* .orange[Correlations:] the average product of the z-score of _x_ and the z-score of _y_

* $r_{x,y} = \frac{1}{n-1} \sum^{n}_{i=1} \big(\frac{x_i - \bar{x}}{S_x} \times \frac{y_i - \bar{y}}{S_y}\big)$

* From -1 to +1, regardless of the original scale

* Correlations only measure _linear_ relationships
]
---

# The Health and Wealth of Nations

```{r}
library(gapminder)
names(gapminder)
cor(gapminder$gdpPercap, gapminder$lifeExp)
cor(gapminder$pop, gapminder$lifeExp)

df <- subset(gapminder, select = c("lifeExp", "pop", "gdpPercap")) # variables
cor(df, use = "pairwise.complete.obs") # pairwise correlations
```

---

# The Health and Wealth of Nations

```{r, tidy=FALSE, cache=TRUE, fig.align='center', warning=FALSE, message=FALSE, fig.width=11, fig.height=5}
plot(gapminder$gdpPercap, gapminder$lifeExp,
     main = "GDP per capita and life expectancy",
     ylab = "Life expectancy",
     xlab = "GDP per capita",
     pch = 16, col = "grey")
```

---

class: inverse, center, middle

# Recap: Log Transformations

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---
# Recap: Log Transformation

```{r, fig.align='center'}
hist(gapminder$gdpPercap, main = "GDP per capita", freq = FALSE)
```
---

# Log Transformation

```{r, fig.align='center'}
hist(log(gapminder$gdpPercap), main = "Log GDP per capita", freq = FALSE)
```

---

# The Health and Wealth of Nations

```{r, tidy=FALSE, cache=TRUE, fig.align='center', warning=FALSE, message=FALSE, fig.width=11, fig.height=5}
plot(gapminder$gdpPercap, gapminder$lifeExp,
     main = "GDP per capita and life expectancy",
     ylab = "life expectancy",
     xlab = "GDP per capita",
     pch = 16, col = "grey")
```
---

# The Health and Wealth of Nations

```{r, tidy=FALSE, cache=TRUE, fig.align='center', warning=FALSE, message=FALSE, fig.width=11, fig.height=4.5}
cor(log(gapminder$gdpPercap), gapminder$lifeExp)
plot(log(gapminder$gdpPercap), gapminder$lifeExp,
     main = "Log GDP per capita and life expectancy",
     ylab = "life expectancy",xlab = "Log GDP per capita",
     pch = 16, col = "grey")
```
---
# Quantile-Quantile Plots

.font150[
* Specific type of correlation

* Compare whole distributions

* Each point represents same quantile

* If 2 distributions are identical, the graph will be a 45-degree line

* `qqplot()`
]
---

# Quantile-Quantile Plots

```{r, tidy=FALSE, cache=TRUE, fig.align='center', warning=FALSE, message=FALSE, fig.height=6}
qqplot(gapminder$gdpPercap, gapminder$lifeExp, main = "QQ plot",
       xlab = "GDP per capita", ylab = "Life expectancy")
```
---

# Quantile-Quantile Plots

```{r, tidy=FALSE, cache=TRUE, fig.align='center', warning=FALSE, message=FALSE, fig.height=6}
qqplot(log(gapminder$gdpPercap), gapminder$lifeExp, main = "QQ plot",
       xlab = "Log GDP per capita", ylab = "Life expectancy")
```
---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 
---

# Matrices

.font150[
* R can perform many operations with matrices

* Data frames have many types of variables (numeric, factor, integer), matrices only take numeric entries

* Select variables: 
  - Data frame: `dataset$variable`
  - Matrices: `matrix[row number,column number]`

* `matrix(numbers, nrow = x, ncol = y, byrow = TRUE/FALSE)`
]
---

# Matrices

```{r}
ma1 <- matrix(1:10, nrow = 2, ncol = 5, byrow = TRUE)
ma1
ma1 * 3
t(ma1)    # transpose
```

---

# Matrices

```{r}
ma1
diag(ma1)            # diagonal
apply(ma1, 1, mean)  # 1 = row, 2 = column
apply(ma1, 1, sum)   # sum of matrix rows
apply(ma1, 2, sd)    # standard deviation of columns
```

---
# Matrices

```{r}
ma2 <- matrix(21:30, nrow = 2, ncol = 5, byrow = FALSE)
ma2

ma1 + ma2  
ma3 <- rbind(ma1, ma2) # concatenate rows, use cbind to aggregate columns
ma3
```
---

# What You Have Learned

.font150[
* Survey sampling: random and quota sampling

* Biases: individual non-response, item non-response, social desirability

* Plots: histograms, bar plots, box plots, scatter plots, time series

* Z-Scores, correlations, quantile-quantile plots, list experiments
]
---

class: inverse, center, middle

# And we are done with measurement!!

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 
---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---

class: inverse, center, middle

# See You next week!

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 