---
title: "Homework 2"
author: "Umberto Mignozzetti"
date: "4/7/2020"
output: pdf_document
---

# Basic R Operations

&nbsp;

## 1.1) Math operations

My first math operation is a summation and a division:

```{r}
(2+2)/4
```

In the first operation, I am summing up two plus two, taking the result, and dividing by four. My second operation is a summation of a log:

```{r}
log(3)+1
```

In the second operation I am taking the natural log of three and summing one to it. My last operation is a product of a difference:

```{r}
3*(pi-1)
```

In the last operation, I am taking the number pi, subtracting one, and multiplying the result by three.

&nbsp;

## 1.2) Install devtools

The following code install packages in R:

```{r}
install.packages('devtools', repos = "http://cran.us.r-project.org")
```

During my installation, I had a problem with an error that was showing up on my knitr. The problem was because I was trying to download a package without setting a mirror. I've searched on the internet and found the [following solution on Stackoverflow](https://stackoverflow.com/questions/33969024/install-packages-fails-in-knitr-document-trying-to-use-cran-without-setting-a). When I added the repos in my install.packages command, all worked fine.

&nbsp;

## 1.3) Installing ggplot2

Looking into the [ggplot2 GitHub page](https://github.com/tidyverse/ggplot2), I've found the following code to install the GitHub version of ggplot2 using devtools:

```{r}
devtools::install_github("tidyverse/ggplot2")
```

And the installations worked well.

&nbsp;

## 1.4) separate function

The first thing I did was to load the package and ask for help in R.

```{r}
library(tidyverse)
help(separate)
```

Then, I googled the functions separate and I've found a webpage explaining how it works. The webpage is in [here](https://tidyr.tidyverse.org/reference/separate.html).

The help in R tells me that the function do the following:

> "Given either regular expression or a vector of character positions, separate() turns a single character column into multiple columns."

&nbsp;

## 1.5) piecharts R

I've googled it and found the [following webpage](https://www.statmethods.net/graphs/pie.html). It has explanations on how to build several types of piecharts. Below I am adding one of the examples in the webpage:

```{r fig.align="center"}
slices <- c(10, 12,4, 16, 8)
lbls <- c("US", "UK", "Australia", "Germany", "France")
pie(slices, labels = lbls, main="Pie Chart of Countries")
```

&nbsp;

## 1.6) apropos function

First, I had no idea what the apropos function was supposed to do. I googled it and find that it searches for functions with names containing a given character expression. Below my answer about which functions have the character test on them:

```{r}
apropos('test')
```

&nbsp;

## 1.7) stackoverflow

I looked into the stackoverflow website and it has many questions and answers related to R and other computer languages. The questions are rated by users, what facilitates to identify which answers are useful.

I took a deeper look into [this question here](https://stackoverflow.com/questions/61076605/how-to-connect-the-point-with-a-line-on-plot-in-r) because it was about plots, and we learned plots during class. The user asked the following:

> "I was trying to plot this dataframe on R and then connect the dots together with a line for having a development graph showing how the number of film per year has changed."

The person wants to build a graph with how the number of films changed per year. The solution was the following:

```{r}
set.seed(111)
fpa = data.frame(Anno=sample(2000:2018),FilmPerAnno=rnbinom(19,mu=10,size=1))
par(mfrow=c(1,2))
plot(fpa, type = "b", xlab="Year", ylab="Film Per Year",main="not ordered")
plot(fpa[order(fpa$Anno),], type = "b", xlab="Year", ylab="Film Per Year",
main="ordered")
```

&nbsp;

## 1.8) my_package

To install the my_package package, we need to use the install.packages function:

```{r eval = FALSE}
install.packages('my_package', repos = "http://cran.us.r-project.org")
```

Once installed, to load the package we need to use the following function:

```{r eval = FALSE}
library(my_package)
```

&nbsp;

## 1.9) my_function

On my_package, I can find about my_function in R by asking for help. I'd use the following syntax:

```{r eval = FALSE}
help(my_function)
```

Outside R, I can take a look on stackoverflow, or even on the creator's website.

&nbsp;

## 1.10) only c() vector

When I put on my console the following code:

```{r eval = FALSE}
myVector <- c()
```

It created a NULL empty vector.

&nbsp;

## 1.11) only c("") vector

When I put on my console the following code:

```{r eval = FALSE}
myVector <- c("")
```

It created a vector with "" inside it.

&nbsp;

## 1.12) combining different objects

The running code for the example is the following:

```{r}
# create vector x
x <- c(1,2,3)

# find the class of x
class(x)

# create vector y
y <- c('1','2','3')

# find the class of y
class(y)

# combining both
z <- c(x,y)

# find the class of z
class(z)
```

It coerces both objects to become character vectors.

&nbsp;

## 1.13) Load mtcars

Looked on the internet and I've found the [following instructions](https://rstudio-pubs-static.s3.amazonaws.com/481654_883a4b47c9b244d4859dd1db235f0165.html) on how to load a built-in R dataset:

```{r}
data(mtcars)
```

To summarize one variable in the dataset, I looked into its head:

```{r}
head(mtcars)
```

And selected the variable mpg:

```{r}
summary(mtcars$mpg)
```

&nbsp;

## 1.14) build table

I build a table of the variable cyl in mtcars:

```{r}
table(mtcars$cyl)
```

&nbsp;

## 1.15) readings

You should do this one by yourself.

&nbsp;

## 1.16) types of variables:

- The chance that it will rain tomorrow: **quant continuous**
- Number of cars that travel, in one day, on the busiest avenue in the city: **quant discrete**
- Customer satisfaction with the service: excellent, good, regular, bad, or very bad: **quali ordinal**
- Blood type: **quali nominal**
- Email of Company X customers: **quali nominal**
- Time of professional experience of employees of Company X, in years: **quanti discrete**
- Time of professional experience of employees of Company X, in seconds: **quanti discrete**
- Professional experience of Company X employees: high experience, medium experience, or low experience: **quali ordinal**
- Gender: **quali nominal**
- HDI of the countries: **quant continuous**
- The inflation rate for the year: **quant continuous**

&nbsp;

# Fix the code

## 2.1)

The original code was this one here:

```{r eval = FALSE}
"2 + 2"
```

And my answer is the following:

```{r}
2 + 2
```

## 2.2)

The original code was this one here:

```{r eval = FALSE}
2 + "pi"
```

And my answer is the following:

```{r}
2 + pi
```

## 2.3)

The original code was this one here:

```{r eval = FALSE}
1 = TRUE
```

And my answer is the following:

```{r}
1 == TRUE
```

## 2.4)

The original code was this one here:

```{r eval = FALSE}
4 \ 3
```

And my answer is the following:

```{r}
4 / 3
```

## 2.5)

The original code was this one here:

```{r eval = FALSE}
Help(cor.test)
```

And my answer is the following:

```{r}
help(cor.test)
```

## 2.6)

The original code was this one here:

```{r eval = FALSE}
???cor.test
```

And my answer is the following:

```{r}
??cor.test
```

## 2.7)

The original code was this one here:

```{r eval = FALSE}
help(if)
```

And my answer is the following:

```{r}
help('if')
```

## 2.8)

The original code was this one here:

```{r eval = FALSE}
?help("cor.test")
```

And my answer is the following:

```{r}
help("cor.test")
```

## 2.9)

The original code was this one here:

```{r eval = FALSE}
this is my sum
# 1 + 2
```

And my answer is the following:

```{r}
# this is my sum
1 + 2
```

## 2.10)

The original code was this one here:

```{r eval = FALSE}
numbers <- c(1 2 3 4)
```

And my answer is the following:

```{r}
numbers <- c(1,2,3,4)
```

## 2.11)

The original code was this one here:

```{r eval = FALSE}
animals <- c("dog" "cat" "parrot" "bunny")
```

And my answer is the following:

```{r}
animals <- c("dog","cat","parrot","bunny")
```

## 2.12)

The original code was this one here:

```{r eval = FALSE}
animals <- C("dog", "cat", "parrot", "bunny")
```

And my answer is the following:

```{r}
animals <- c("dog", "cat", "parrot", "bunny")
```

## 2.13)

The original code was this one here:

```{r eval = FALSE}
Head(mtcars)
```

And my answer is the following:

```{r}
head('mtcars')
```

## 2.14)

The original code was this one here:

```{r eval = FALSE}
str(Mtcars)
```

And my answer is the following:

```{r}
str(mtcars)
```

## 2.15)

The original code was this one here:

```{r eval = FALSE}
dim[mtcars]
```

And my answer is the following:

```{r}
dim(mtcars)
```

## 2.16)

The original code was this one here:

```{r eval = FALSE}
nomes(mtcars)
```

And my answer is the following:

```{r}
names(mtcars)
```

## 2.17)

The original code was this one here:

```{r eval = FALSE}
head(mtcars, x = 10)
```

And my answer is the following:

```{r}
head(mtcars, n = 10)
```

## 2.18)

The original code was this one here:

```{r eval = FALSE}
animals <- c("dog", "cat", "parrot", "bunny")
numbers <- c(1,2,3,4,5,6,7,8,9,10)
dat <- data.frame(animals,numbers)
```

And my answer is the following:

```{r}
animals <- c("dog", "cat", "parrot", "bunny", NA, NA, NA, NA, NA, NA)
numbers <- c(1,2,3,4,5,6,7,8,9,10)
dat <- data.frame(animals,numbers)
dat
```

## 2.19)

The original code was this one here:

```{r eval = FALSE}
dat <- c(1:100)
save("my_data.RData", file = dat)
```

And my answer is the following:

```{r}
dat <- c(1:100)
save(dat, file = "my_data.RData")
```

## 2.20)

The original code was this one here:

```{r eval = FALSE}
data("diamonds")
ggplot(data = aes(x = cut)) +
  geom_bar()
```

And my answer is the following:

```{r}
data("diamonds")
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

## 2.21)

The original code was this one here:

```{r eval = FALSE}
data("diamonds")
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = cut))
```

And my answer is the following:

```{r}
data("diamonds")
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

## 2.22)

The original code was this one here:

```{r eval = FALSE}
theme_bw() +
ggplot(data = diamonds, mapping = aes(x = cut))
```

And my answer is the following:

```{r}
ggplot(data = diamonds, mapping = aes(x = cut)) +
  theme_bw()
```

## 2.23)

The original code was this one here:

```{r eval = FALSE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, color = "red"))
```

And my answer is the following:

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut), color = "red")
```

# Challenge

My vector is the following:

```{r}
x <- 1:1000000
```

I have no idea on how to do this! But the hint says:

> "Hint: we don’t want you to type each element. Find out how to select elements in a vector. Hint: the remainder of the division can be obtained by using the command %%. For instance, type 5 %% 3 in R"

Test: if I had a vector y with four elements:

```{r}
y <- c(1,2,3,4)
```

The even numbers are 2 and 4. To select them:

```{r}
y[c(2,4)]
```

And using remainders:

```{r}
y[y %% 2 == 0]
```

And for the entire vector:

```{r}
head(x[x%%2 == 0], 20)
```