---
title: "Week 05a - Measurement, Missing Observations, and Data Visualization"
subtitle: "Plots and Survey Sampling <html><div style='float:left'></div><hr color='#EB811B' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: "03/24/2020"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "macros.js"
      highlightStyle: github
      ratio: 4:3
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "http://cran.cnr.berkeley.edu/"
options(repos = r)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #EB811B;
}

.orange {
  color: #EB811B;
}
</style>

# Today's Agenda

.font150[
* Tables: `table()` and `prop.table()` 

* Missing data in R: `is.na()` and `na.omit()`

* Visualizing data: `barplot()`, `hist()`, and `boxplot()`

* Survey sampling methods 
]

---

class: inverse, center, middle

# Tables

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---

# Tables in R

.font130[
* We have learned how to use the `table()` command in R

* Example:
]

.code120[
```{r tidy=FALSE}
afghan <- read.csv("https://raw.githubusercontent.com/umbertomig/intro-prob-stat-FGV/master/datasets/afghan.csv")
names(afghan)
table(ISAF = afghan$violent.exp.ISAF,
      Taliban = afghan$violent.exp.taliban)
```
]
---

# Tables in R: prop.table()

.font130[
* We can also use `prop.table()` to see the proportion of cases in each cell

* We have to include `table()` within parentheses too:
]

.code120[
```{r tidy=FALSE}
prop.table(table(ISAF = afghan$violent.exp.ISAF,
                 Taliban = afghan$violent.exp.taliban))
```
]
---
# Tables in R

.font130[
* Since we're already using nested functions, we can also use `round()` to round the values in each cell

* Notice the `, 2` in the code below. It indicates that we will round the numbers up to two significant digits
]

.code120[
```{r tidy=FALSE}
round(prop.table(table(ISAF = afghan$violent.exp.ISAF,
                       Taliban = afghan$violent.exp.taliban)), 2)
```
]

.font150[
* The table is now much easier to read and it conveys the same information
]
---

# Tables in R: prop.table()

.code120[
```{r tidy=FALSE}
round(prop.table(table(ISAF = afghan$violent.exp.ISAF,
                       Taliban = afghan$violent.exp.taliban)), 2)
```
]

.font150[
* What's the percentage of people that have been victimized by the Taliban?

* And by both the Taliban and ISAF?
]
---

# Tables in R: prop.table()

.font150[
* Another example, now with 3 significant digits and multiplied by 100
]

.code120[
```{r tidy=FALSE}
round(prop.table(table(Employed = afghan$employed,
                       Income = afghan$income)), 3) * 100
```
]
---


class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---

class: inverse, center, middle

# Missing data

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---

# Missing Data

.font150[
* Not all individuals answer to surveys

* Two types of non-response:
  - Individual non-response
  - Item non-response
  
* Both tend to bias the results

* So it is very important that we know where (and think about why) we see gaps in our data
]
---

# Missing Data

.font150[
* R has a special code for missing data, `NA`

* Since `NA` is only used for missing observations, we can count their numbers with `is.na()`
]

---

# Missing Data

.code120[
```{r}
head(afghan$income, n = 10)
sum(is.na(afghan$income)) # number of missings
round(mean(is.na(afghan$income)), 2) # proportion of missings
```
]
---

# Missing Data

.font150[
* Some R function don't work if there's missing data

* We add `na.rm = TRUE` to the code
]

.code120[
```{r}
# Victims of Taliban violence
sum(is.na(afghan$violent.exp.taliban))
mean(afghan$violent.exp.taliban)
round(mean(afghan$violent.exp.taliban, na.rm = TRUE), 2)
```
]

.font150[
* Why do you think we have missing data here?
]

---

# Missing Data

.font150[
* We can create a new variable or data set without missing observations

* Use the `na.omit()` command
]

.code150[
```{r}
length(afghan$violent.exp.taliban)
taliban.no.missing <- na.omit(afghan$violent.exp.taliban)
length(taliban.no.missing)
length(afghan$violent.exp.taliban) - length(taliban.no.missing)
```
]

---

# Missing Data

.font150[
* You can also visualize the number of missing observations with `summary()`
]

.code150[
```{r}
summary(afghan$violent.exp.taliban)
sum(is.na(afghan$violent.exp.taliban))
```
]
---

class: inverse, center, middle

# Data Visualization

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 
---

# Bar Plots

.font120[
* Bar plots are used to visualize .orange[factor/character variables]

* Proportion of observations in each category as the height of each bar

* Options:

  - `main = "Title"`
  - `xlab = "X label"`
  - `ylab = "Y label"`
  - `xlim = c(number, number)` limits for the x variable  
  - `ylim = c(number, number)` limits for the y variable  
  - `names.arg = c("Bars labels")` - in the same order of the variable
  - `horiz = TRUE` for horizontal plots
  - `cols = "colour name"` bar color (see: )
  
* You can use `barplot()` with `prop.table()` instead of pie charts
]
---

# Bar Plots

```{r}
employed.ptable <- prop.table(table(afghan$employed))
employed.ptable
```

```{r tidy=FALSE, eval=FALSE}
employed.ptable <- prop.table(table(afghan$employed))
barplot(employed.ptable,
        names.arg = c("Unemployed", "Employed"), 
        main = "Proportion of Employed Afghanis",
        xlab = "Employment",
        ylab = "Proportion",
        ylim = c(0, 0.6))
```
---

# Bar Plots

```{r tidy=FALSE, echo=FALSE, fig.align='center'}
barplot(employed.ptable,
        names.arg = c("Unemployed", "Employed"), # 0 and 1, respectively
        main = "Proportion of Employed Afghanis",
        xlab = "Employment",
        ylab = "Proportion",
        ylim = c(0, 0.6))
```

---

# Bar Plots

```{r tidy=FALSE, eval=FALSE}
barplot(employed.ptable,
        names.arg = c("Unemployed", "Employed"), # 0 and 1, respectively
        main = "Proportion of Employed Afghanis",
        ylab = "Employment", # change the axes
        xlab = "Proportion", 
        xlim = c(0, 0.7), # now it's xlim 
        horiz = TRUE,     # because the plot is horizontal
        col = "brown")
```
---

# Bar Plots

```{r tidy=FALSE, echo=FALSE, fig.align='center'}
barplot(employed.ptable,
        names.arg = c("Unemployed", "Employed"), # 0 and 1, respectively
        main = "Proportion of Employed Afghanis",
        ylab = "Employment",
        xlab = "Proportion",
        xlim = c(0, 0.6),
        horiz = TRUE,
        col = "brown")
```

---


class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---

class: inverse, center, middle

# Histograms

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---

# Histograms

.font120[
* We use histograms to display the distribution of a .orange[numeric variable]

* They are similar to bar plots

* Numeric variables are _binned_ into groups

* Histograms shows the density of each bin

* Important: Height is share of observations in bin divided by bin size

* We care less about the density of each bin than about the distribution of the variable as a whole

* Area of each bar is the share of observations that fall into that bin

* Area of all bins sum to one
]
---

# Histograms

.font120[
* Many options are similar to those of `barplot()`: `main`, `xlab`, `ylim`, `col`

* We can also add `freq = FALSE` to show the density of each histograms

* `breaks =` changes the size of the bins

* Densities are useful to compare different distributions

* .orange[Densities are not percentages]: "percentage per horizontal unit"
]
---

# Histograms

```{r tidy=FALSE, eval=FALSE}
# For colours, see: http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf
hist(afghan$age,
     main = "Histogram - Age",
     xlab = "Age",
     xlim = c(0, 0.04),
     freq = FALSE,
     col = "darkorange2") 
```
---

# Histograms

```{r tidy=FALSE, echo=FALSE, fig.align='center'}
hist(afghan$age,
     main = "Histogram - Age",
     xlab = "Age",
     ylim = c(0, 0.04),
     freq = FALSE,
     col = "darkorange2") 
```
---

# Histograms

.font150[
* We can also add text and fitted lines to all R plots

* Use `text()` and `abline()` after `hist()`
]

```{r tidy=FALSE, eval=FALSE}
hist(afghan$age,
     main = "Histogram - Age",
     xlab = "Age",
     ylim = c(0, 0.04),
     freq = FALSE,
     col = "darkorange2") 
## add a text label at (x, y) = (35, 0.35)
text(x = 35, y = 0.035, "median")
## add a vertical line representing median
abline(v = median(afghan$age))
```
---

# Histograms

```{r tidy=FALSE, echo=FALSE, fig.align='center'}
hist(afghan$age,
     main = "Histogram - Age",
     xlab = "Age",
     ylim = c(0, 0.04),
     freq = FALSE,
     col = "darkorange2") 
## add a text label at (x, y) = (35, 0.35)
text(x = 35, y = 0.035, "median")
## add a vertical line representing median
abline(v = median(afghan$age))
```

---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---

class: inverse, center, middle

# Box-Plots

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---
# Box Plots

.font150[
* Like histograms, box plots also display the distribution of a numeric variable

* Box plots show _the median_, _quartiles_, and _IQR_

* Useful to compare different distributions side-by-side

* It is also useful to identify outliers, that is, data points that are above 1.5 times the interquartile range (IQR)

* `boxplot()`

]

---

# Box Plots

.font150[
* `boxplot()` also has a series of optional arguments:

  - `main`, `ylab`, `ylim`, `col`
  - `formula = y ~ group`, y is numeric variable and group is a factor
]

---
# Box Plots

```{r tidy=FALSE}
names(afghan)
median(afghan$age)
quantile(afghan$age, probs = c(0, .25, .5, .75, 1))
```
---

# Box Plots

```{r tidy=FALSE, fig.align='center'}
boxplot(afghan$age, main = "Distribution of age", ylab = "Age", ylim = c(10, 80))
```
---

# Box Plots and Histograms

.font150[
* To visualize the distribution of a single variable, histograms might be more informative. Compare:
]

```{r tidy=FALSE, eval=FALSE}
hist(afghan$age,
     main = "Histogram - Age",
     xlab = "Age",
     ylim = c(0, 0.04),
     freq = FALSE) 
## add a text label at (x, y) = (35, 0.35)
text(x = 35, y = 0.035, "median")
## add a vertical line representing median
abline(v = median(afghan$age))
```
---

# Histogram

```{r tidy=FALSE, echo=FALSE, fig.align='center'}
hist(afghan$age,
     main = "Histogram - Age",
     xlab = "Age",
     ylim = c(0, 0.04),
     freq = FALSE) 
## add a text label at (x, y) = (35, 0.35)
text(x = 35, y = 0.035, "median")
## add a vertical line representing median
abline(v = median(afghan$age))
```
---

# Box Plots

```{r tidy=FALSE, fig.align='center'}
boxplot(afghan$age, main = "Age Distribution", ylab = "Age", ylim = c(10, 80))
```
---

# Box Plots

.font150[
* But box plots provide an easy way to compare multiple observations at a time

* Similar to `tapply()`
]

```{r}
tapply(afghan$age, afghan$province, median, na.rm = TRUE)
```
---

# Box Plots

```{r tidy=FALSE, fig.align='center'}
boxplot(afghan$age ~ afghan$province,
        main = "Age Distribution by Province", ylab = "Age", col = "grey")
```

---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---

class: inverse, center, middle

# Survey Sampling

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---

# Survey Sampling

.font130[Last primaries:]

.pull-left[
#### Democratic candidates

* Hillary Clinton 44%
* Bernie Sanders 26% 
* Joe Biden 20%
]

.pull-right[
#### Republican candidates

* Donald Trump 26%
* Ben Carson 15%
* Carly Fiorina 10%
* Jeb Bush 9%
* Marco Rubio 9%
* Ted Cruz 6%
]

.font110[Source: Huffington Post, Pollster]

.font130[
* Most polls only interview several hundred voters
]
--
.font130[
* Goal: infer what 200 million voters are thinking
]
---

# The 1936 Literary Digest Poll

.font150[
* Mail questionnaire to 10 million people

* Final sample size: over 2.3 million returned

* Addresses came from phone books and club memberships

* The young Gallup used 50,000 respondents
]


.center[
|                 | FDR's vote share | 
|:----------------|-----------------:|
| Literary Digest |               43 | 
| George Gallup   |               56 | 
| Actual Outcome  |               62 |
]
---

# The 1936 Literary Digest Poll

.font150[
* Why did the Literary Digest get the numbers so wrong?
]
--
.font150[
* .orange[Biased sample]
]

---

# Quota Sampling vs Random Sampling

.font150[
* **Quota sampling:** Sample certain groups until quota is filled

* Problem: Unobservables can bias the results
]
--
.font150[
* **Random sampling:** Random draws without replacement from the population

* Everybody has the same chance of being in the sample

* Problem: _none_, sample is unbiased! 
]
---

# Random Sampling

.font150[
* Not every single sample will match all characteristics of the population exactly

* But as the number of samples gets larger (say 1000 samples of 1000 respondents), on average the samples would be representative

* Polls are associated with _uncertainty_: plus or minus a number

* But getting a random sample is _hard_
]

---

# Difficulties

.font130[
* Problems of telephone survey
  - Random digit dialing from phone book
  - Wealthy individuals have higher changes of being called
  - Caller ID screening (unit non-response)

* Problems of internet survey
  - Non-probability sampling
  - Cheap but non-representative
  - Young, urban, rich groups are overrepresented
  - Requires statistical corrections (usually weights)
]

---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 
---

# Homework

.font150[
* Changing minds about gay marriage?

* You are going to perform a replication of an experiment.

* It is going to be amazing!
]

---

class: inverse, center, middle

# See you next Thursday!

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 
